---
title: "Homework_12"
author: "Noah Beckage"
date: "4/13/2022"
output: html_document
---

[Home page](index.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE)
```

#### Libraries
```{r libraries}
library(ggplot2)
library(RColorBrewer)
library(viridisLite)
library(paletteer)
library(ggpubr)
library(grid)
```


# Read in Data
This data set contains 10 molecular properties from a random sample of 88,878 chemicals from the PubChem database. First column `SMILES` contains a unique <a href="https://en.wikipedia.org/wiki/Simplified_molecular-input_line-entry_system" target="_blank">smile string</a> for each chemical. There is also a column with G-protein coupled receptor (GPCR) activity scores and one with qualitative estimates of drug-likeness (QED).

For ease of computing, I'm going to take a random sample of 10,000 rows from this data frame.

```{r read data}
all_data <- read.csv("C:/Users/nbeck/OneDrive/Desktop/Spring 2022/CS 287/Project/CS287-Project/data/PubChemData.csv")

small_data <- all_data[sample(1:nrow(all_data),10000),]
str(small_data)
```

# How does molecular weight relate to GPCR activity scores?
```{r make plot functions}
thermal <- colorRampPalette(c("black","blue", "#007FFF", "cyan",
                                 "#7FFF7F", "yellow", "#FF7F00", "red", 
                                 "white"))
  
##################################################
# FUNCTION: MW_dplot -> Molecular weight Density Plot
# Makes nice density plots of the specified molecular property vs GPCR Activity
# input: df->source data frame; x(str)->col to use for x value; y(str)-> col to use for y values; palette(str)->name of palette to be used (must be in viridus palette); xlimit(int/double)->x limit for the graph; the rest of the input are strings specifying graph labels
# output: p -> prints the plot
#------------------------------------------------- 
MW_dplot <- function(df=small_data,
                           x='MW',
                           y='GPCR_act',
                           palette="magma",
                           xlimit=NULL,
                           xlab='Molecular Weight',
                           ylab='GPCR Activity Score',
                           title=NULL,
                           subtitle=NULL,
                           cap=NULL,
                           legend='Density') {
  p <- ggplot(data=df,mapping=aes_string(x=x,y=y))+
  stat_density_2d(geom = "raster", aes(fill = stat(density)), contour = FALSE)+
  scale_fill_viridis_c(option=palette)+
  theme_classic()+
  labs(x=xlab,y=ylab,fill=legend)
  if(!is.null(xlimit)){
    p <- p+xlim(0,xlimit)
  }
  if(!is.null(title)){
    p <- p+labs(title=title)
  }
  if(!is.null(subtitle)){
    p <- p+labs(subtitle=subtitle)
  }
  if(!is.null(cap)){
    p <- p+labs(caption=cap)
  }
 return(p)
}
```

```{r make plots}
mw_inferno <- MW_dplot(palette = 'inferno',xlimit=700,xlab=NULL,ylab=NULL,title='Inferno')

mw_magma <- MW_dplot(palette = 'magma',xlimit=700,xlab=NULL,ylab=NULL,title='Magma')

mw_plasma <- MW_dplot(palette = 'plasma',xlimit=700,xlab=NULL,ylab=NULL,title='Plasma')

mw_thermal <- MW_dplot(xlimit=700,xlab=NULL,ylab=NULL,title='Thermal (custom)')+
  scale_fill_gradientn(colors=thermal(6))
```

```{r make figure}
figure <- ggarrange(mw_inferno,mw_magma,mw_plasma,mw_thermal,ncol=2,nrow=2)
figure <- annotate_figure(figure, left = textGrob("GPCR Activity", rot = 90, vjust = 1, gp = gpar(cex = 1.5)),
                    bottom = textGrob("Molecular Weight", gp = gpar(cex = 1.5)))
figure
# figure %>% ggexport(filename = "hw12_LARGE_plot.pdf")
```
After many hours, *that* is a figure I can proudly display. The last line I commented out creates a <a href="hw12_LARGE_plot.pdf" target="_blank">pdf</a> of the plot that would be ready for publication. That pdf is generated from the full dataset (n=88,878), but for this R markdown I'm running the smaller dataset (n=10,000) to save some computing time. It looks pretty, but all in all I don't think this would be the most accurate visualization of these data. The intense color scales tend to obscure the sample size of the data and over-emphasize density; the density function and intense colors also mask outliers (from looking at a scatter plot, I know there were a few very heavy chemicals that don't show up well in these plots). I've also been told that the density function uses a gaussian kernel, and although I don't totally understand what that means, I know that we can't assume these data are gaussian. 

All in all though, I do find these plots satisfying. I'll say that between all these weedy ggplot-based packages and Python's matplotlib (which my lab tends to use for visualizations), making nice graphs, let alone figures, can easily become a headache. All the more reason this was a useful exercise.

