---
title: "Homework_11"
author: "Noah Beckage"
date: "4/6/2022"
output: html_document
---

SCROLL TO BOTTOM TO SEE WHATS LEFT TO DO

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

[Home page](index.html)

```{r}
library(dplyr)
```


### Question 1
```{r}
##################################################
# function: file_builder
# create a set of random files for regression
# input: file_n = number of files to create
#       : file_folder = name of folder for random files
#       : file_size = c(min,max) number of rows in file
#       : file_na = number on average of NA values per column
# output: set of random files
#------------------------------------------------- 
file_builder <- function(file_n=10,
                        file_folder="Data/RandomFiles/",
                        file_size=c(15,100),
                        file_na=3){
for (i in seq_len(file_n)) {
file_length <- sample(file_size[1]:file_size[2],size=1) # get number of rows
var_x <- runif(file_length,max=0.75) # create random x with predetermined max
var_y <- runif(file_length,min=0.25) # create random y with predetermined min
df <- data.frame(var_x,var_y) # bind into a data frame
bad_vals <- rpois(n=1,lambda=file_na) # determine NA number
df[sample(nrow(df),size=bad_vals),1] <- NA # random NA in var_x
df[sample(nrow(df),size=bad_vals),2] <- NA # random NA in var_y

# create label for file name with padded zeroes
file_label <- paste(file_folder,
                       "ranFile",
                       formatC(i,
                       width=3,
                       format="d",
                       flag="0"),
                       ".csv",sep="")

# set up data file and incorporate time stamp and minimal metadata
write.table(cat("# Simulated random data file for batch processing","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# NBB","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=file_label,
                    row.names="",
                    col.names="",
                    sep=""))

# now add the data frame
write.table(x=df,
            file=file_label,
            sep=",",
            row.names=FALSE,
            append=TRUE)


}
}
```

```{r}
##################################################
# function: t_stats
# performs t test, returns p value
# input: 2-column data frame (x and y)
# output: p-value
#------------------------------------------------- 
t_stats <- function(d=NULL) {
             if(is.null(d)) {
               x_var <- runif(10,max=0.75)
               y_var <- runif(10,min=0.25)
               d <- data.frame(x_var,y_var)
             }
  t <- t.test(x=d[,1],y=d[,2],alternative='less')
  return(t$p.value)

}
```


```{r results='hide'}
#--------------------------------------------
# Global variables
file_folder <- "Data/RandomFiles/"
n_files <- 100
file_out <- "Data/StatsSummary.csv"
#--------------------------------------------

# Create 100 random data sets
dir.create(file_folder)
file_builder(file_n=n_files)
file_names <- list.files(path=file_folder)

# Create data frame to hold file summary statistics
ID <- seq_along(file_names)
file_name <- file_names
slope <- rep(NA,n_files)
p_val <- rep(NA,n_files)
r2 <- rep(NA,n_files)

stats_out <- data.frame(ID,file_name,p_val)

# batch process by looping through individual files
for (i in seq_along(file_names)) {
  data <- read.table(file=paste(file_folder,file_names[i],sep=""),
                     sep=",",
                     header=TRUE) # read in next data file
  
  d_clean <- data[complete.cases(data),] # get clean cases
  
  . <- t_stats(d_clean) # pull regression stats from clean file
  stats_out[i,3] <- . # copy into last column
  
}
# set up output file and incorporate time stamp and minimal metadata
  write.table(cat("# Summary stats for batch processing of t-tests","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# NBB","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=file_out,
                    row.names="",
                    col.names="",
                    sep=""))
  
# now add the data frame
  write.table(x=stats_out,
              file=file_out,
              row.names=FALSE,
              col.names=TRUE,
              sep=",",
              append=TRUE)
```

```{r}
df <- read.table("Data/StatsSummary.csv",sep=',',header = TRUE)
head(df)
```


### Question 2
(Continuation of Homework 10)
```{r}
# Initializing a vector with some random number of zeroes
vect=sample(c(-10:10),size=100,replace=T)

# This line of code returns the number of zeroes in that vector
length(subset(vect,vect==0))
```

### Question 3
```{r}
##################################################
# FUNCTION: product_matrix
# creates a matrix where each element is the product of its respective row # and col #
# input: nrows, ncols -> number of rows and cols (ints)
# output: mat -> matrix of size nrows, ncols
#------------------------------------------------- 
product_matrix <- function(nrows=runif(1,min=1,max=10),ncols=runif(1,min=1,max=10)){
  mat=matrix(data=NA, nrow=nrows,ncol=ncols)
  if(nrows<=0 | ncols<=0){
    return(message("Error: nrows and ncols must both be greater than zero"))
  }
  for(i in 1:nrow(mat)){
    for(j in 1:ncol(mat)){
      mat[i,j] <- i*j
      }
  }
 return(mat)
}
product_matrix()
```

### Question 4
#### Function to read in (or generate) the data
```{r}
# can't forget to set seed
set.seed(42)

####### ###########################################
# function: read_data
# read in (or generate) data set for analysis
# input: file name (or nothing, for this demo)
# output: 3 column data frame of observed data (x,y)
#------------------------------------------------- 
read_data <- function(z=NULL) {
  if(is.null(z)){
    # random group assignment
    DO_CHEMISTRY <- sample(c('YES','NO'),20,T)
    # minimum MW is 15 (methane), max chosen arbitrarily
    MW <- sample(15:1000,20,F)
    z <- data.frame(INDEX=1:20,DO_CHEMISTRY,MW)} 
  else {
    z <-read.table(file=z,header=TRUE,sep=",")}
  return(z)
}
```

#### Function to calculate the means for each treatment group
```{r}
##################################################
# function: get_metric
# perform t-test for means of the two treatment groups for randomization test
# input: 3-column data frame for t-test
# output: p-value from t-test
#------------------------------------------------- 
get_means <- function(z=NULL) {
  if(is.null(z)){
    # random group assignment
    DO_CHEMISTRY <- sample(c('YES','NO'),20,T)
    # minimum MW is 15 (methane), max chosen arbitrarily
    MW <- sample(15:1000,20,F)
    z <- data.frame(INDEX=1:20,DO_CHEMISTRY,MW)} # set up data frame                 
    group_means <- z %>% group_by(z$DO_CHEMISTRY) %>% 
    summarise(mean=mean(z$MW))
  # Row names bug needs to be addressed
  t <- data.frame(t(group_means))[-1,]
  names(t) <- c("NO","YES")
  return(t)
}
get_means(df)
```

#### Function to perform t-test and return p-value
```{r}
##################################################
# function: get_pval
# calculate p value from simulation
# input: dataframe of randomized (or real) data
# output: t-test p-value
#------------------------------------------------- 
get_pval <- function(z=NULL) {
  if(is.null(z)){
    z <- data.frame(NO=sample(15:1000,20),YES=sample(15:1000,20))}
  else{
    t <- pivot_wider(z,id_cols=CPD_ID,names_from=DO_CHEMISTRY,values_from=MW)}
  pval <- t.test(z[,"YES"],z[,"NO"])$p.value
return(pval)
}
get_pval()
```
#### Function to randomize the data
```{r}
##################################################
# function: shuffle_data
# randomize data for regression analysis
# input: 2-column data frame (x_var,y_var)
# output: 2-column data frame (x_var,y_var)
#------------------------------------------------- 
shuffle_data <- function(z=NULL) {
  if(is.null(z)){
    # random group assignment
    DO_CHEMISTRY <- sample(c('YES','NO'),20,T)
    # minimum MW is 15 (methane), max chosen arbitrarily
    MW <- sample(15:1000,20,F)
    z <- data.frame(INDEX=1:20,DO_CHEMISTRY,MW)}
  z[,2] <- sample(z[,2])
  return(z)
}
```

#### Putting it all together in program body

Your functions should all be working now. Double-check that the text in the comments are correct after you finish. Ok, you have to get a p-value from a randomized, "shuffled" data set. Then compare that to the p-value from the regular (non-shuffled). Run a short loop (100x) to do 100x randomization tests and generate 100 pvalues. Now compare the average of those 100 randomized pvalues to your regular pvalue. Still different? Explain why.DONE.

```{r}
shuffle_data(read_data("Data/NIHMS50746-supplement-AZ_chemsurvey_QED.csv") %>% 
  select(CPD_ID,DO_CHEMISTRY, MW))


```

